\section{Future Work}

While the Darwin MAC Framework and the SEDarwin policy module are
usable in the current prototype, further research and development
is necessary.  The following list summarizes the significant remaining
areas of development:

\begin{itemize}

\item {\bf Update source tree to Mac OS 10.4.X ``Tiger''} - The Framework and
SEDarwin module are currently based on Apple's Mac OS 10.3.8 ``Panther''
release.  The Darwin MAC Framework has already been ported to Tiger,
however, the SEDarwin module and some Mach-specific Framework changes
still remain to be ported.

\item {\bf Update SEDarwin security server} - The FLASK security
server and AVC components should be updated to include changes that
have been made on SELinux.

\item {\bf Increase entry point coverage} - The SEDarwin policy module
does not yet implement all entry points provided by the Darwin MAC
Framework.

\item {\bf Finer granularity of Mach IPC controls} - While we currently
have coverage for Mach message send and receive as well as port
rights transfer and receipt the existing entrypoints do not provide
the same degree of control that was present in DTOS.

\item {\bf Better Mach kernel object coverage} - We currently only
label Mach ports and tasks.  Coverage should be extended to Mach
memory objects, semaphores, timer objects, swap controls and the
Mach messages themselves.

\item {\bf Distributed file systems} - Additional work is necessary to
explore the impact of mandatory protections and labeling on
distributed file systems.

\item {\bf Review system boot} - A more thorough examination of the
boot process is necessary, including implications for labeled security
policies.

\item {\bf Installation} - It is still cumbersome to install and
configure a system to run the MAC Framework and labeled policies.
Additional work is necessary to simplify the installation process.

\item {\bf IOKit} - In order to identify kernel services offered by
device drivers that require additional labeling and access control,
the IOKit device driver extension framework needs to be reviewed.

\item {\bf Investigate System Services} - Review some critical system
services, such as, kextd, security server, netinfo, and configd for
security implications.

\item {\bf SEBSD/SEDarwin policy} - The TE policy/rules for FreeBSD
and Darwin are rather immature; they should be updated to include
relevant changes from SELinux system, as well as custimizations for
the FreeBSD/Darwin operating environments.

\item {\bf Improved test coverage} - The test framework has only begun
to test elements of labeled security policies and should be extended.

\end{itemize}
